---
title: "Hower_DataViz_Lab01"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### **Load Packages**
  + load *here*, *tidyverse* and *extrafont* for additional functionality


```{r}
library(here)
library(tidyverse)
library(extrafont)
library(wesanderson)
```
  
### **Assign data to variable**
  + read data from csv file and assign to object called "hot_dogs"


```{r}
setwd( "C:/Users/User/Documents/DataViz/")
hot_dogs <- read_csv(here::here("hot_dog_contest.csv"),col_types = cols(
  gender = col_factor(levels = NULL)
))
glimpse(hot_dogs)
```


### **Wrangle Data**  
  + filter for males after 1981 and create a boolean vector indicating IFOCE years
  
  
```{r} 
hot_dogs <- hot_dogs %>%
  filter(gender == "male", year > 1981) %>%
  mutate(postIFOCE = year >= 1997)
```

### **Base Plot**


```{r}
ggplot(hot_dogs, aes(x = year, y = num_eaten))+
geom_col()
```


### **CHALLENGE 1**
### Plot 1
  + add title and white borders around bars


```{r}
ggplot(hot_dogs, aes(x = year, y = num_eaten))+
geom_col(colour = "white") + 
labs(x ="year", y ="number eaten") +
ggtitle("Hot Dog Eating Contest Results, 1981 - 2017")
```


### Plot 2
  + color bars navy blue


```{r}
ggplot(hot_dogs, aes(x = year, y = num_eaten))+
geom_col(colour = "white", fill = "navyblue") + 
labs(x ="year", y ="number eaten") +
ggtitle("Hot Dog Eating Contest Results, 1981 - 2017")
```

### Plot 3
  + color bars by IFOCE years to make plot multivariate


```{r}
ggplot(hot_dogs, aes(x = year, y = num_eaten))+
geom_col(colour = "white", aes(fill = postIFOCE)) + 
labs(x ="year", y ="number eaten") +
ggtitle("Hot Dog Eating Contest Results, 1981 - 2017")
```


### Final Plot
  + Reformat color legend labling to improve readability
  
```{r}
ggplot(hot_dogs, aes(x = year, y = num_eaten))+
geom_col(aes(fill = postIFOCE), colour = "white") + 
labs(x ="year", y ="number eaten") +
ggtitle("Hot Dog Eating Contest Results, 1981 - 2017") +
theme(legend.title=element_blank()) +
scale_fill_discrete(name = "",
                      labels=c("Pre-IFOCE", "Post-IFOCE"))
``` 


### Bonus Plot
  + change lightness and chroma as an experiment
  
```{r}
ggplot(hot_dogs, aes(x = year, y = num_eaten))+
geom_col(aes(fill = postIFOCE), colour = "white") + 
labs(x ="year", y ="number eaten") +
ggtitle("Hot Dog Eating Contest Results, 1981 - 2017") +
scale_color_manual("legend", values = wes_palette("FantasticFox"))+
theme(legend.title=element_blank()) 

```




